---
title: "Statistical Inference Project"
author: "ST Lin"
date: "April 13, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
```

## Part One

Investigate the exponential distribution in R and compare it with the Central Limit Theorem.
**lambda** is the rate parameter for the exponential distribution.
The **mean** of exponential distribution is 1/lambda and the **standard deviation** is also 1/lambda.

- Set lambda = 0.2 for all of the simulations. 

- Investigate the distribution of averages of 40 exponentials.

- Do a thousand simulations.

Set the simulation parameters.
```{r}
lambda <- 0.2
n_trial <- 40
n <- 1000
```

###1. Show the sample mean and compare it to the theoretical mean of the distribution.
Compute the mean and standard deviation of theoretical exponential distribution.
The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda.
```{r}
meanExp <- 1/lambda
sdExp <- (1/lambda)/sqrt(n_trial)
meanExp; sdExp
```
Distribution of simulated means of exponential distribution
```{r}
set.seed(333)
mns = NULL
for (i in 1 : n) mns = c(mns, mean(rexp(n_trial, lambda)))
meanSample <- mean(mns)
sdSample <- sd(mns)
meanSample; sdSample
```
The means and standard deviations of the exponential and sample mean distributions (1,000 samples) are close to each other.

###2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
```{r}
varExp <- sdExp^2
varSample <- var(mns)
varExp; varSample
```

Plot theoretical exponential distribution and sample mean distribution. They look very different. The latter takes on a bell shape distribution.
```{r}
par(mfrow=c(1,2))
hist(rexp(n, lambda), breaks=20, main="Exponential distribution", xlab="")
hist(mns, breaks=20, main="Sample means distribution", xlab="")
```


###3. Show that the distribution is approximately normal.
```{r}
par(mfrow=c(1,1))
hist(mns, freq=FALSE, breaks=20, main=NULL, xlab = "mean of sampled exponential distribution")
curve(dnorm(x,mean=meanExp,sd=sdExp),from=0, to=10, col='red',add=T) 
```

By visual examination, the sample mean distribution approximates a normal distribution of mean = meanExp and standard deviation = sdExp.

##Part Two
###1. Load the ToothGrowth data and perform some basic exploratory data analyses.
```{r}
dataTG <- (ToothGrowth)
dim(dataTG)
summary(dataTG)
str(dataTG)
head(dataTG)
table(dataTG$supp,dataTG$dose)
```
###2. Provide a basic summary of the data.
There are 2 supp types: VC and OJ, each with 30 records of len values given dose being 0.5, 1.0 or 2.0.
Generate boxplots to visually examine the distribution of len by supp and dose.(ggplot2 returns some warnings, which will not affect the analysis.)
```{r}
library(ggplot2)
dataTG$dose <- as.factor(dataTG$dose)
g <- ggplot(dataTG, aes(x=dose, y=len, fill=dose))
g <- g + geom_boxplot()
g <- g + facet_grid(.~supp)
g <- g + ggtitle("Distribution of tooth length over supp and dose")
g
```

##3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering.)
Conduct t test over supp type
```{r}
gOJ <- dataTG[dataTG$supp == "OJ",]
gVC <- dataTG[dataTG$supp == "VC",]
t.test(gOJ$len, gVC$len, paired = FALSE)
```
The 95 percent confidence interval is: (-0.1710156  7.5710156). We **cannot reject the null hypothesis that the difference of lengths between VC and OJ is zero**.

Perform t test on OJ and VC groups over different doses.

###OJ
```{r}
gOJ05 <- gOJ[gOJ$dose == 0.5,]
gOJ10 <- gOJ[gOJ$dose == 1.0,]
gOJ20 <- gOJ[gOJ$dose == 2.0,]
```

Pairwise t tests over doses, given OJ.

- Dose 0.5 vs. 1.0
```{r}
t.test(gOJ05$len, gOJ10$len, paired = FALSE)
```
Reject null hypothesis that dose change has no effect on length, given 95% confidence interval.

- Dose 1.0 vs. 2.0
```{r}
t.test(gOJ10$len, gOJ20$len, paired = FALSE)
```
Reject null hypothesis that dose change has no effect on length, given 95% confidence interval.

- Dose 0.5 vs. 2.0
```{r}
t.test(gOJ05$len, gOJ20$len, paired = FALSE)
```
Reject null hypothesis that dose change has no effect on length, given 95% confidence interval.

All 3 t tests indicate we can **reject null hypothesis that dose does not have effect on length, given supp is OJ**.

###VC
```{r}
gVC05 <- gVC[gVC$dose == 0.5,]
gVC10 <- gVC[gVC$dose == 1.0,]
gVC20 <- gVC[gVC$dose == 2.0,]
```
Pairwise t tests over doses, given VC.

- Dose 0.5 vs. 1.0
```{r}
t.test(gVC05$len, gVC10$len, paired = FALSE)
```
Reject null hypothesis that dose change has no effect on length, given 95% confidence interval.

- Dose 1.0 vs. 2.0
```{r}
t.test(gVC10$len, gVC20$len, paired = FALSE)
```
Reject null hypothesis that dose change has no effect on length, given 95% confidence interval.

- Dose 0.5 vs. 2.0
```{r}
t.test(gVC05$len, gVC20$len, paired = FALSE)
```
Reject null hypothesis that dose change has no effect on length, given 95% confidence interval.

All 3 t tests indicate we can **reject null hypothesis that does does not have effect on length, given supp is VC**.

##4. State your conclusions and the assumptions needed for your conclusions.
By t tests along supp, there is no significance of difference in effects of OJ and VC. However, the dose does have significant effect on th length, be supp VC or OJ.
