---
title: "markdown_demo"
author: "Fred"
date: "March 30, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an **R Markdown document**. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see *<http://rmarkdown.rstudio.com>*.

### Checking your R environment

```{r env}
sessionInfo()
```

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

```{r}
table(cars)
```
## You can create a table.
```{r table}
library(datasets)
library(xtable)
fit <- lm(Ozone ~ Wind + Temp + Solar.R, data = airquality)
xt <- xtable(summary(fit))
print(xt, type = "html")
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
